Finding a closest point in a lattice of Voronoi's first kind

### 2014-07-08 reviews/response.tex

\newcommand{\rng}{\operatorname{rng}}
\newcommand{\subrng}{\operatorname{subr}}
\newcommand{\decrng}{\operatorname{decrng}}

\item\textbf{Comment:} 
Almost all the inequalities in the paper can be derived from the following simple obser-
vation.  For and $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
\[
\xbf^T L_G \ybf = \sum_{ \{i,j\} \in E[G] } \wbf_{ij} (\xbf_{i} - \xbf_{j}) (\ybf_{i} - \ybf_{j}) \geq 0.
\]
Furthermore, if $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \geq \dots \geq \ybf_{\pi[n]}$ then the above inequality if reversed, i.e $\xbf^T L_G \ybf \leq 0$.

I believe using the above will make many of your proofs more transparent.  For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then noice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order. In particular this implies that
\[
\xbf L_G \xbf = (\xbf - 1_S)^T L_G (\xbf - 1_S) + 2(\xbf - 1_S)^T L_G 1_S + 1_S^T L_G 1_S \geq (\xbf - 1_S)^T L_G (\xbf - 1_S),
\]
i.e. $\xbf$ can be made ``shorter'' by squeezing its components closer together.

Furthermore, one can use this to give a two line proof of Theorem 4.1.  This theorem I actually do think make sense to reprove here, because modifications of this proof are what drives the convergence bound. 
\\\textbf{Response:}
BLERG. This is indeed the property used to prove Theorem 4.1, but not what's needed for the convergence

@@

\end{enumerate}

\section*{Reviewer 2}\label{sec:reviewer-2}

\begin{enumerate}
\item\textbf{Comment:}  
page 1, last line "The closest lattice point problem is known to be NP-hard under certain condition ...": I found this paragraph very confusing, and somehow inaccurate. What "certain conditions"? I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following. CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance. Interestingly (see 35 and 17) there are families of lattices when CVP remain NP-hard even when given the best possible basis (or any hard-to-compute side information about the lattice). This version of the problem is called CVP with preprocessing (CVPP), and it is particularly relevant to this work, where it is shown that CVPP can be solved exactly and in polynomial time for the class of lattices of Voronoi first kind, when given as auxiliary information a special basis (namely, an obtuse superbasis.) 
\\\textbf{Response:}

\item\textbf{Comment:}  
Additional comments about the above paragraph: one author is missing in the reference for 12. Of [1,23,40] only 23 achieves $n^{O(n)}$ running time. (I think the algorithms in 40 take $2^{O(n^2)}$.) There is a journal version of 37 in SIAM J. Computing, and the paper is by two authors (not just Micciancio). 
\\\textbf{Response:}
We have fixed these errors.  This paragraph now reads:

``The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.''

@@

\item\textbf{Comment:}  
page 3, prop 2.2, again definition of ceiling(r) is non-standard. Use floor(r)+1 instead. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 3, section 3. 
The process described in (3.1) is almost identical to those in [45,37], so those papers should probably be referenced at the very beginning, and a comparison provided. As far as I remember, the difference is that 37 picks the v that minimizes the voronoi norm (i.e., the norm defined by the voronoi cell of the origin) instead of the euclidean norm. On the other hand, 45 uses the euclidean norm, but selects an arbitrary relevant vector that reduces the euclidean norm, rather than the one that reduces it the most. In this respect, your process is a special case of the one described in 45, and termination (your propositions 3.1 and 3.2) is already proved in 45. 
\\\textbf{Response:}
We have included the following text in the first paragraph of Section~\zref{sec:iterative-slicer} when introducing the interative algorithm.

``Motivated by Shalvi~et.~al.~\cite{Shalvi_iterativeslicer_2009} and  Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013}, we consider the following iteration,''

Among the iterative proceedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.  The choice of our iteration is not motivated by what might be fastest in general lattices.  It is motivated by what we can prove to be polynomial time for those lattices of Vornoi's first kind.

The~\emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009} is similar but uses the (presumably accelerated?) iteration
\begin{align*}
x_{k+1} &= x_k + \alpha_k v_k  \\
(v_k, \alpha_k) &= \arg\min_{ (v, \alpha) \in ( \relevant(\Lambda) \cup \{\zerobf\}, \ints) } \|y - x_k - \alpha v \|.
\end{align*} 
So it allows adding integer multiples of a relevant vector.  We do not believe using this will improve the complexity of our algorithm beyond $O(n^4)$.  It would greatly complicate our proofs.

The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.  To our (admittedly rough) knowledge, the proceedure computes an $\alpha_k$ (different, but related to that above) and then proceeds by the same iteration as ours to compute a closest point in the scaled lattice $\alpha_k \Lambda$.  The next $\alpha_{k+1}$ is computed and the proceedure repeats.

We beleive that detailing the differences between these iterative proceedures would only detract from the core purpose of our paper to decscribe an $O(n^4)$ time algorithm to compute a closest point in lattice of Vornoi's first kind.

@@

- D. Aharonov and Oded Regev (2005). Lattice problems in NP ∩ co-NP. J. ACM 52, 749-765.

(Gives polynomial time solution to approximate the distance estimation version of CVPP within $\gamma=\sqrt(n \log n)$ approximation factor, i.e., for any lattice there is an appropriate advice (more complex than just a good basis) that allows to approximate in polynomial time the distance of any target point to the lattice within gamma, but without actually finding a nearby lattice point.)

- On the Closest Vector Problem with a Distance Guarantee 
D. Dadush, O. Regev, N. Stephens-Davidowitz, Computational Complexity Conference 2014, to appear (next week!)

I am not sure, as the paper has not appeared yet, but I suspect it is the same as the paper on Dadush webpage with title "On BDD and CVPP". 
See link 
http://cs.nyu.edu/~dadush/papers/cvpp.pdf 
and read abstract for more information.
\\\textbf{Response:}
We thank the reviewer for these interesting references highlighting the difficultly of computing close approximations of a closest lattice point.  These have been referenced in Section~\zref{sec:iterative-slicer} with the following text:

``Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aharonov_Regev_2005,Aleknovish_hardness_with_preprocessing_2011,Dadush_cvp_with_distance_guarantee_2014}.''

### 2014-07-08 paper.tex
more comments.

\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson}

@@

The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

@@

The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.  Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $(\floor{r/\rho}+1)^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.  Thus, the number of iterations required might grow exponentially with $n$.  The number of iterations required depends on the lattice point that starts the iteration $x_0$.  It is helpful for $x_0$ to be, in some sense, a close approximation of the closest point $x_K$.  Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing}.  We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.  Combining this with the fact that each iteration of the algorithm requires $O(n^3)$ operations results in an algorithm that requires $O(n^4)$ operations to compute a closest point in a lattice of Voronoi's first kind.

@@

An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  That is, there exists a set of $n+1$ vectors $b_1,\dots,b_{n+1}$ such that $b_1,\dots,b_n$ are a basis for $\Lambda$,
\begin{equation}\label{eq:superbasecond}
b_1 + b_2 \dots + b_{n+1} = 0
\end{equation}
(the \emph{superbasis} condition), and the inner products satisfy
\begin{equation}\label{eq:obtusecond}
q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
\end{equation}
(the \emph{obtuse} condition).  The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.  It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.

@@

@@

We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  We want to find $n$ integers $w_1,\dots,w_n$ that minimise
\[
\| y - \sum_{i=1}^n b_i w_i \|^2.
\]
We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
\[
\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
\]
The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.  In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.  This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.  Let
\begin{equation}\label{eq:matrxBobtusebasis}
B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
\end{equation}
be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.  We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
\begin{equation}\label{eq:tominimise}
\| B(z  -  w) \|^2
\end{equation}
is minimised.  Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ operates on vectors elementwise. In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
\begin{align}
x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
u_{k+1} &= u_k + t_k \nonumber \\
t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
\end{align}
where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.  The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.  This choice of initial lattice point is important.  In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  The minimiser may not be unique corresponding with the existence of multiple minimum cuts.  In this case any one of the minimisers may be chosen.  Our results do not depend on this choice.   In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.  That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.

@@

\begin{lemma}\label{lem:decrngpreservesclosestpoints}
If $v \in \ints^{n+1}$ such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
\end{lemma}
\begin{proof}
The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.  It remains to prove the lemma when $\subrng(v) \neq \emptyset$.  In this case put $S = \subrng(v)$ and put 
\[
u = \decrng(v) = v - \onebf_S.
\] 
Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.  We have $\zeta - u = \zeta - v + \onebf_S$.  Applying part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = \zeta - v$ we obtain
\begin{align}
\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \Phi(S, \zeta-v) \nonumber \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j)\big). \label{eq:sumsumBzBu}
\end{align}
Observe that $\zeta_i =  z_i - \floor{z_i} \in [0,1)$ for all $i=1,\dots,n+1$ and so $-1 < \zeta_i-\zeta_j < 1$ for all $i,j=1,\dots,n+1$.  Also, for $i \in S$ and $j\notin S$ we have 
\[
v_i - v_j \geq \min\{ v_i, i \in S\} - \max\{v_j, j \notin S\} \geq 2
\]
by definition of $\subrng(v) = S$.  Thus,
\[
1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}.
\]
Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ for $i \neq j$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
\[
\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0.
\]
It follows that $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
\end{proof}

>>>>

\author{Robby~G.~McKilliam, Alex~Grant and I.~Vaughan~L.~Clarkson}

@@

The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard,Regev_2004_inappox_lattice_with_preprocessing,feige_inapproximability_2004,Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

@@

The other factor affecting the complexity is the number of iterations required before the algorithm arrives at a closest lattice point, that is, the size of $K$.  Proposition~\ref{eq:latticepointsinsphere} suggests that this number might be as large as $(\floor{r/\rho}+1)^n$ where $r = \|y - x_0\|^2$ and $\rho$ is the packing radius of the lattice.  Thus, the number of iterations required might grow exponentially with $n$.  The number of iterations required depends on the lattice point that starts the iteration $x_0$.  It is helpful for $x_0$ to be, in some sense, a close approximation of the closest point $x_K$.  Unfortunately, computing close approximations of a closest lattice point is known to be computationally difficult~\cite{feige_inapproximability_2004,Regev_2004_inappox_lattice_with_preprocessing,Aharonov_Regev_2005,Aleknovish_hardness_with_preprocessing_2011,Dadush_cvp_with_distance_guarantee_2014}.  We will show that for a lattice of Voronoi's first kind a simple and easy to compute choice for $x_0$ ensures that a closest lattice point is reached in at most $n$ iterations and so $K \leq n$.  Combining this with the fact that each iteration of the algorithm requires $O(n^3)$ operations results in an algorithm that requires $O(n^4)$ operations to compute a closest point in a lattice of Voronoi's first kind.

@@

An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  That is, there exists a set of $n+1$ vectors $b_1,\dots,b_{n+1}$ such that $b_1,\dots,b_n$ are a basis for $\Lambda$,
\begin{equation}\label{eq:superbasecond}
b_1 + b_2 \dots + b_{n+1} = 0
\end{equation}
(the \emph{superbasis} condition), and the inner products satisfy
\begin{equation}\label{eq:obtusecond}
q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
\end{equation}
(the \emph{obtuse} condition).  The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.  It is known that all lattices in dimensions less than or equal to $3$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  
%BLERG.  THIS MIGHT BETTER GO IN THE GRAPH SECTION
% There is a one-to-one correspondence between the obtuse superbasis of a lattice of Vornoi's first kind and what is called the \emph{Laplacian} of a weighted graph.  To se this, let
% \[
% B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
% \]
% be the matrix with columns given by $b_1,\dots,b_{n+1}$ and define the $n+1$ by $n+1$ \emph{Gram matrix} $Q = B^\prime B$.  The elements of $Q$ are given by the Selling parameters, that is, $Q_{ij} = q_{ij}$.  Let $G$ be a graph with $n+1$ vertices $v_1,\dots,v_{n+1}$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.  To each edge assign the \emph{weight} $w_{ij} = -q_{ij} \leq 0$.  Let $d_{ii} = \sum_{j \neq i} w_{ij}$ be the sum of weights of edges connected to vertex $v_i$. The Laplacian of the graph $G$ is the $n+1$ by $n+1$ matrix with elements
% \[
% L_{ij} = \begin{cases}
% d_{ii} & i = j \\
% -w_{ij} & i \neq j. 
% \end{cases}
% \]
An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.

@@

BLERG: It will be instructive for us to give a proof of this theorem.  In doing so we will introduce some of the notation that will be used in Section~\ref{}.

@@

We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  We want to find $n$ integers $w_1,\dots,w_n$ that minimise
\[
\| y - \sum_{i=1}^n b_i w_i \|^2.
\]
We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
\[
\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
\]
The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.  In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.  This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a the closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.  Let
\[
B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
\]
be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.  We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
\begin{equation}\label{eq:tominimise}
\| B(z  -  w) \|^2
\end{equation}
is minimised.  Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ operates on vectors elementwise. In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
\begin{align}
x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
u_{k+1} &= u_k + t_k \nonumber \\
t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
\end{align}
where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.  The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.  This choice of initial lattice point is important.  In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  The minimiser may not be unique corresponding with the existence of multiple minimum cuts.  In this case any one of the minimisers may be chosen.  Our results do not depend on this choice.   In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.  That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.

@@

\begin{lemma}\label{lem:decrngpreservesclosestpoints}
If $v \in \ints^{n+1}$ is such that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$, then $B\big(\floor{z} + \decrng(v)\big)$ is also a closest lattice point to $y$.
\end{lemma}
\begin{proof}
The lemma is trivial if $\subrng(v) = \emptyset$ so that $\decrng(v) = v$.  It remains to prove the lemma when $\subrng(v) \neq \emptyset$.  In this case put $S = \subrng(v)$ and put 
\[
u = \decrng(v) = v - \onebf_S.
\] 
Let $\zeta = z - \floor{z}$ be the column vector containing the fractional parts of the elements of $z$.  We have $\zeta - u = \zeta - v + \onebf_S$.  Applying part~\ref{eq:lem:decSellingsinc} of Lemma~\ref{lem:decSellings} with $p = \zeta - v$ we obtain
\begin{align}
\|B(\zeta - v)\|^2 - \|B(\zeta - u)\|^2 &= \Phi(S, \zeta-v) \nonumber \\
&= \sum_{i \in S}\sum_{j \notin S}q_{ij}\big(1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j)\big). \label{eq:sumsumBzBu}
\end{align}
Observe that $\zeta_i =  z_i - \floor{z_i} \in [0,1)$ for all $i=1,\dots,n+1$ and so $-1 < \zeta_i-\zeta_j < 1$ for all $i,j=1,\dots,n+1$.  Also, for $i \in S$ and $j\notin S$ we have 
\[
v_i - v_j \geq \min\{ v_i, i \in S\} - \max\{v_j, j \notin S\} \geq 2
\]
by definition of $\subrng(v) = S$.  Thus,
\[
1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}.
\]
Substituting this inequality into~\eqref{eq:sumsumBzBu} and using that $q_{ij} \leq 0$ for $i \neq j$ (the obtuse condition~\eqref{eq:obtusecond}) we find that
\[
\|B(z - \floor{z} - v)\|^2 - \|B(z - \floor{z} - u)\|^2 \geq 0.
\]
It follows that $B(\floor{z} + u) = B\big(\floor{z} + \decrng(v)\big)$ is a closest lattice point to $y = Bz$ whenever $B(\floor{z} + v)$ is.
\end{proof}
